Developing a preprint culture in biology

### 2013-02-26 ms.tex

The ongoing discussions on the scientific publication process, peer-review and
alternative publication models are all symptoms of the current uneasiness with
the ever growing obsession with bibliographic metrics such as the impact factor
\cite{Fisher2012}. Researchers are pressured to orient their publication
strategy to maximize their number of publications and total citations. A
well-known consequence is the initial submission of manuscripts to the most
prestigious journals, followed by resubmission to progressively ``lower level''
journals as they are rejected. The numerous negative impacts of such behavior
have been discussed in depth \cite{hoc09} and include a long delay between the
time a manuscript is finished and its publication. Research activities and the
publication process are drifting away from their fundamental objective, namely
the distribution of novel scientific discoveries.

Developing a preprint culture in biology will not solve all of the problems with
the current publication process. However, it might significantly reduce the
negative consequences. The role of peer-review is to judge the scientific
quality of a study. It is the first barrier against the fraudulent and poor
quality science that could impede scientific progress. In practice, the
peer-review system is not only used to evaluate scientific quality but also to
judge pertinence. On the other hand, preprints are not filtered on either
quality or pertinence. Widespread adoption of preprint servers has the potential
to shift the distribution strategy: journals would remain important to validate
publications, but the relevance of a study should only be judged by many more
readers than the typical two to four anonymous reviewers. With a shift in the
diffusion strategy, the role of traditional journals and their editors would be
to showcase scientific discoveries for specialized readership.

>>>>

The ongoing discussions on the scientific publication process, peer-review and
alternative publication models are all symptoms of the current uneasiness with
the ever growing obsession with bibliographic metrics such as the impact factor
\cite{Fisher2012}. Researchers are pressured to orient their publication
strategy to maximize their number of publications in top journals and total citations. A
well-known consequence is the initial submission of manuscripts to the most
prestigious journals, followed by resubmission to progressively ``lower level''
journals as they are rejected. The numerous negative impacts of such behavior
have been discussed in depth \cite{hoc09} and include a long delay between the
time a manuscript is finished and its publication. Research activities and the
publication process are drifting away from their fundamental objective, namely
the distribution of novel scientific discoveries.

Developing a preprint culture in biology will not solve all of the problems with
the current publication process. However, it might significantly reduce the
negative consequences. The role of peer-review is to judge the scientific
quality of a study. It is the first barrier against fraudulent and poor quality
science that could impede scientific progress. In practice, the peer-review
system is not only used to evaluate scientific quality but also to judge
pertinence. In contrast, preprints are not filtered based on either quality or
pertinence. Widespread adoption of preprint servers has the potential to shift
the distribution strategy: journals would remain important for validating
publications, but the relevance of a study should only be judged by many more
readers than the typical two to four anonymous reviewers. With a shift in the
diffusion strategy, the role of traditional journals and their editors would be
to showcase scientific discoveries for specialized readership.
